### 2015-09-24 exos/exosText_MIT.tex
added Matts stuff but still need to get his bib files

% % Created 2015-09-15 Tue 11:46
% \documentclass[11pt]{article}
% \usepackage[utf8]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage{fixltx2e}
% \usepackage{graphicx}
% \usepackage{longtable}
% \usepackage{float}
% \usepackage{wrapfig}
% \usepackage{rotating}
% \usepackage[normalem]{ulem}
% \usepackage{amsmath}
% \usepackage{textcomp}
% \usepackage{marvosym}
% \usepackage{wasysym}
% \usepackage{amssymb}
% \usepackage{hyperref}
% \usepackage{color}
% \usepackage{soul}
% \tolerance=1000
% \usepackage[margin=1in]{geometry}

% \newcommand{\hilight}[1]{\colorbox{yellow}{#1}}

% %\author{Alex Ansari}
% %\date{}
% %\title{TALOS}
% %\hypersetup{
% %  pdfkeywords={},
% %  pdfsubject={},
% %  pdfcreator={Emacs 24.3.1 (Org mode 8.2.10)}}

% \begin{document}
\section{MIT Exoskeleton}

Design is based  on fact that evidence from biology and passive walking devices suggest that legged locomotion can be very efficient.  The main physical concept behind this efficiency being that there is a gait cycle for a pair of legs that naturally exchanges energy between gravity and inertia.  The main mechanical design principles centered around using a combination of passive and active elements.  In particular, the MIT design uses {\bf springs, variable impedance joints, and powered actuators.}  The suit generally  functions as a lightweight, underactuated robot which runs in parallel to an operator and supports the weight of a payload.

Additionally, the leg structure of the suit allowed weight from the suit as well as payload to be transferred directly to the ground.

Additionally, there are several other design principles that the MIT team discusses which are especially relevant to this report.  The first is that joint {\bf powers scale linearly with mass}  The second is that altering the gait pattern of the operator has been shown to increase the physiological energy expended during locomotion.  Thus, the specifications for actuation and control for their system are extracted from the angle, torque , and power data of human walking joint patterns.

\subsection{Actuator specifications}

The MIT exoskeleton uses a single actuator located at the hip.  The actuator is designed around a total system weight of 165 kg.  The maximum scaled hip torque for a system with this mass is approximately negative 130 Nm during the stance phase and approximately 100 Nm during the swing phase of the gait.  The hip is chosen as the actuated joint because proximal mass is metabolically less expensive in walking than distal mass.

The specific choice of actuation for the MIT exoskeleton hip joint was a {\bf linear series elastic actuator}.  This choice was made because it provided a lightweight and relatively inexpensive means of implementing force control with a bandwidth similar to that of natural muscle.  A model of the actuator is shown in Figure \ref{fig:mitSEA}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/mitSEA}
  \caption{}
  \vspace{-0.2in}
 \label{fig:mitSEA}   
 \end{figure} 
Like all series elastic designs the actuator shown in Figure \ref{fig:mitSEA} includes a spring in series with the output of the motor drive system of the actuator.  The motor-drive mechanism consists of a {\bf brushed DC motor driving a 3mm ball screw via a 2:1 reduction belt drive}.  The nut of the ball screw is coupled to the actuator output shaft via four compression die springs.  The compression of the die springs are measured using a linear potentiometer, which is how forces at the output are measured.

The motor chosen for the {\bf actuator was a 150W Maxon RE40 Brushed DC motor}, which was chosen based on the bio-mechanical data and for its power and weight ratio.

The force bandwidth for the actuator was characterized during both stance and swing phases of the walking gait cycle.  {\bf During stance, the bandwidth of the actuator was 35 Hz, while the bandwidth during swing was found to be 40Hz.}

 \subsection{Exoskeleton Specifications}

 The MIT exoskeleton design has a total of fourteen degrees of freedom, three at each hip, one at each knee, two at each ankle, and one in each foot.

 A cam mechanism was implemented at the hip joint to enable hip ab/adduction.  This system was necessary because, during abduction, there is was a length difference between the operator and exoskeleton leg which resulted from having non-collocated centers of rotation.  The cam mechanism automatically adjusts the exoskeleton leg length such that the center of rotation of the exoskeleton hip is projected onto the biological hip center of rotation.  A picture of the MIT exoskeleton is shown in Figure \ref{fig:MITsuit}.
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/MITsuit}
  \caption{}
%  \vspace{-0.2in}
 \label{fig:MITsuit}   
 \end{figure}

 The interface between the suit and operator consisted of shoulder straps, a compliant waist belt, thigh cuffs, and a shoe connection.

 The {\bf variable damper knee mechanism} in the flexion/extension degree of freedom was implemented using a magnetorheological damper.  The damper could exert a {\bf maximum braking torque of 60 Nm and consumed on average 1W of electrical power}.

 A linear spring located at the ankle was designed to capture negative energy during dorsilflexion.  The energy stored was released during plantar flexion to add energy to the system as the foot comes off the ground. A urethane spring was used as a liner spring in a lever compression assembly.  The spring constant of the urethane was 247 kN/m, the lever arm of the assembly was 0.0381 m, which gave the rotary spring a stiffness of 356 Nm/rad.

 A unidirectional spring assembly was also used in the hip ab/adduction degree of freedom.  The spring allowed the hip to freely abduct.  It was calculated that 8 Nm were required during adduction.  To this end, the effective rotary stiffness of the spring assembly was 96 Nm/rad.

The suit was powered using a 48V battery pack.

The suit was instrumented with rotary potentiometers at the hip and knee.  Additionally, strain gauges were placed on the structure of the exoskeleton shank, which measured sagittal bending moment and compression force

 \subsection{Control Specifications}

 The control of the MIT exoskeleton functions around the fact that desired actuation and active damping at the knee are functions of gait cycle, at that the values of these desired functions could be determined from human walking data.  Fundamentally, the control strategy works as a state machine, which uses joint angle and measured forces to implement state transitions.

\begin{figure}[thpb]
\centering
\includegraphics[width=4.in]{exos/figs/MIT/walkingCycle}
  \caption{}
  \vspace{-0.2in}
 \label{fig:walkingCycle} 
 \end{figure} 
The walking cycle around which the MIT control scheme was designed is shown in Figure \ref{fig:walkingCycle}.  The individual control strategies in the various phases of the walking cycle are highlighted in Figure \ref{fig:walkingCycle}.  For the hip there are four distinct strategies.  During the thrust phase, the hip exerts a torque which helps to raise the center of mass of the system.  During the extension spring phase, a virtual spring stiffness is programmed which allows energy to be virtually stored while the center of mass moves forward.  During the swing assist phase the virtual energy is released, resulting in a torque being applied which assists in swinging the leg forward.  Finally, in leg retraction a torque is applied to help with foot placement and weight acceptance.  A diagram which illustrates how the sensors in the shank and hip joint are used to switch between these various phases is show in Figure \ref{fig:hipControl} 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/hipControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:hipControl} 
 \end{figure}

 The knee controller in the MIT exoskeleton effectively functions independently from he hip controller.  During knee strike, the variable damper in the knee exerts a torque proportional to the rotational velocity of the knee joint.  A residual magnetic field remained in the knee joint after it was turned off, creating a resistive torque.  The knee needed to be actively demagnetized at full extension during the last phase of stance to allow the knee to swing freely during the swing phase.  The state machine diagram that illustrates how the various phase transitions for the knee joint were implemented using the exoskeleton's onboard sensors is shown in Figure \ref{fig:kneeControl} 
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/kneeControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:kneeControl} 
 \end{figure}

%  \begin{figure}[thpb]
%\centering
%\includegraphics[width=3.in]{figs/torqueCon}
%  \caption{}
%  \vspace{-0.2in}
% \label{fig:IHMCTOR}   
% \end{figure}
%

 \subsection{Opinion}

 The MIT hardware and control system present a very well designed and thought out system which focus primarily on efficiency in design.  The series elastic linear ball screw hardware design seems to be a very smart way to implement efficient drives as well as to perform high fidelity force control.  Using a strain gauge in the shank was also a novel means of of measuring ground reaction forces.  This design may offer benefits in terms of repeatability over different operators as foot fall patterns for different users may include significant variations.

 The details on the controllers were somewhat sparse in the sense that, at least in \cite{}, the exact means through which the added torques were derived was not provided.  Overall though, the controllers were intelligently designed in the sense that they explicitly took the biomechanics of walking into account in their design and implementation.  A similar approach for a variety of different behaviors may be very beneficial to the TALOS project.

% \end{document}

% Conor James Walsh and Daniel Paluska and Kenneth Pasch and William Grand and Andrew Valiente and Hugh Herr, "Development of a lightweight, underactuated exoskeleton for load-carrying augmentation," Proceedings of the 2006 IEEE International Conference on Robotics and Automation, 2006, pp. 3485 - 3491.

% Conor James Walsh and Kenneth Pasch and Hugh Herr, "An autonomous, underactuated exoskeleton for load- carrying augmentation, " International Conference on Intelligent Robots and Systems, 2006, pp. 1410-1415.