### 2015-09-30 survey-atkeson.tex
more updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Appendix: Other exoskeleton systems}

\subsection{Overviews}

\url{http://powerexoskeletons.com/}\\
\url{http://exoskeletonreport.com/}\\
\url{http://robohub.org/tag/exoskeleton/}\\
\url{http://robohub.org/exoskeletons-new-and-older/}\\
\url{http://neurogadget.com/tag/exoskeleton}\\
\url{http://www.engadget.com/tag/exoskeleton/}\\
\url{http://spectrum.ieee.org/robotics/medical-robots/exoskeletons-around-the-world}\\
\url{http://spectrum.ieee.org/biomedical/bionics/the-rise-of-the-body-bots}\\
\url{http://www.extremetech.com/electronics/139633-will-we-ever-have-iron-man-exoskeletons}\\
\url{prezi.com/jygcvfxiesnm/a-brief-introduction-to-biomechanical-exoskeletons/}\\
\url{http://news.discovery.com/tech/robotics/exoskeleton-robots-top-5.htm}\\
\url{http://blog.equipoisinc.com/new-technologies-emerging-aerospace-defense-manufacturing/}\\
\url{http://nextbigfuture.com/2015/03/lower-body-exoskeleton-audi-chairless.html}\\
\url{http://www.army-technology.com/features/featuremilitary-exoskeletons-uncovered-ironman-suits-a-concrete-possibility}\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We do not have sufficient information to survey the following exoskeletons.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Commercial exoskeletons for augmentation}

\subsubsection{Sarcos XOS/WEAR}

\noindent
We do not have sufficient non-proprietary information to survey
this system.\\
\url{http://www.robaid.com/bionics/raytheons-second-generation-exoskeleton-xos-2.htm}\\
\url{http://techcrunch.com/2010/09/27/sarcos-xos2-the-real-life-iron-man/}\\

\subsubsection{Lockheed Martin HULC, FORTIS}

\noindent
We believe this is in collaboration with Ekso Bionics and is similar
to BLEEX.\\
\url{http://www.lockheedmartin.com/us/products/exoskeleton/hulc.html}\\
\url{http://www.lockheedmartin.com/us/products/exoskeleton/FORTIS.html}\\
\url{http://www.lockheedmartin.com/us/news/features/2014/mfc-103114-relief-daily-grind-industrial-exoskeletons-work.html}\\
\url{http://www.dailytech.com/From+HULC+to+FORTIS+the+Evolution+of+Lockheed+Martins+Incredible+Exosuit/article36421.htm}\\
\url{http://www.wired.com/2014/09/navys-exoskeleton-could-make-workers-20-times-more-productive/}\\
\url{http://robrady.com/design-project/lockheed-martin-fortis-human-powered-exoskeleton}\\
\url{http://www.cnn.com/2013/05/22/tech/innovation/exoskeleton-robot-suit/}

\subsubsection{Revision Military}

\noindent
PROWLER Human Augmentation System (HAS) Exoskeleton.\\
\url{https://www.youtube.com/watch?v=RKcqHaPhkkM}\\
\url{https://www.youtube.com/watch?v=Zq4amM9u-6o}\\

\subsubsection{Activelink/Panasonic}

\noindent
A variety of exoskeleton products.\\
\url{http://activelink.co.jp/en}\\

Assist Suit AWN-03:\\
{\it consists of a 13-pound backpack outfitted with leg supports which allows its wearer to lift up to 33 pounds without any noticeable effort.}\\
\url{http://www.engadget.com/2015/08/20/japans-top-oil-company-is-building-an-aliens-power-loader/}\\

Power Loader Suit:\\
\url{http://pinktentacle.com/2009/09/power-loader-exoskeleton-suit/}\\
\url{http://www.extremetech.com/extreme/173463-worlds-first-affordable-powered-exoskeleton-is-almost-here-prepare-for-mech-wars}\\
\url{http://www.dealstreetasia.com/stories/panasonic-to-release-commercial-exoskeleton-to-market-in-4q-2015-9269/}\\

\subsubsection{DAEWOO}

\noindent
\url{https://www.newscientist.com/article/mg22329803.900-robotic-suit-gives-shipyard-workers-super-strength#.U9_FTPldWSp}\\
\url{http://www.geek.com/science/daewoo-dock-workers-now-using-strength-enhancing-exo-skeletons-1601290/}\\

\subsubsection{Innophys: Muscle Suit}

\noindent
Appears to be a power assist suit for lifting. Sip and puff internface?
Tokyo University of Science Prof. Hiroshi Kobayash.\\
\url{https://innophys.jp/}\\
video: \url{https://www.youtube.com/watch?v=Sqr1l909WPw}\\
video: \url{https://www.youtube.com/watch?v=mNdJgh0ZcrM}\\
video: \url{https://www.youtube.com/watch?v=NtX81pf-3G4}\\
\url{http://blogs.wsj.com/japanrealtime/2014/11/12/wearable-power-assist-device-goes-on-sale-in-japan/}

\subsubsection{Noonee (Swiss)}

\noindent
Chair exoskeleton. Passive?\\
\url{http://www.extremetech.com/extreme/188417-rejoice-commuters-and-workers-the-chairless-chair-exoskeleton-lets-you-sit-down-anywhere-anytime}\\

\subsubsection{Sagawa Electronics (Swiss)}

\noindent
Increases reach. Master exoskeleton inside larger slave exoskeleton. Running
demonstrated.\\
\url{https://www.youtube.com/watch?v=tYgHdxmCcAI}\\
\url{http://www.ohgizmo.com/2013/07/11/mk3-exoskeleton-suit-promises-to-make-schoolgirls-better-taller-stronger/}\\
\url{http://www.bitrebels.com/technology/powered-jacket-exoskeleton-japan/}\\

\subsubsection{RB-3D}

\noindent
Hercule\\
\url{http://www.rb3d.com/produits/exosquelettes/}\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Academic exoskeletons for augmentation}

\subsubsection{Kanagawa Institute of Technology}

\noindent
Pneumatic exoskeleton developed at the Kanagawa Institute of Technology,
in Atsugi, Japan, allows Akiko Michihisa, a fitness trainer, ...\\
\url{http://www.ubergizmo.com/2007/10/air-pressure-exoskeleton/}\\

\subsubsection{Monash University/Chen}

\noindent
\url{http://www.dailymail.co.uk/sciencetech/article-2639939/Super-fireman-The-hi-tech-exoskeleton-firefighters-superhuman-abilities.html}\\
\url{http://news.discovery.com/tech/robotics/firefighter-exoskeleton-to-the-rescue-140521.htm}\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Commercial exoskeletons for assistance and rehabilitation}

\subsubsection{ReWalk}

\noindent
On the sixth generation of this device.\\
\url{http://rewalk.com/}\\

\subsubsection{Cyberdyne}

\noindent
up to HAL 5 (already covered)\\
\url{http://spectrum.ieee.org/robotics/medical-robots/exoskeletons-are-on-the-march}\\
\url{http://www.computerweekly.com/photostory/2240108388/Photos-Cyberdyne-Hal-robotic-exoskeleton-to-help-paralyzed/7/Japan-beats-the-US-to-it-Cyberdyne-Hal-robotic-exoskeleton-to-help-paralyzed}\\
\url{http://www.dailymail.co.uk/sciencetech/article-2384930/Robotic-exoskeleton-help-rehabilitate-disabled-people-passes-safety-tests--paving-way-sale-UK.html}\\

\subsubsection{Rexbionics}

\noindent
REX exoskeleton\\
\url{http://www.rexbionics.com/}\\
\url{http://www.gizmag.com/rex-robotic-exoskeleton/15736/pictures}\\
\url{http://www.stripes.com/news/exoskeleton-could-benefit-troops-with-spinal-cord-injuries-1.113657}\\
\url{http://www.funis2cool.com/cool/rex-the-robotic-exoskeleton-could-change-the-lives-of-disabled-people.html}\\

\subsubsection{Indego/Parker Hannafin}

\noindent
Commercialization of Vanderbilt exoskeleton\\
\url{http://www.indego.com/indego/en/home}\\
\url{https://www.youtube.com/watch?v=-bYYmZNxaZk}\\
\url{http://www.roboticstrends.com/article/parker_hannifin_indego_exoskeleton}\\
\url{http://www.crainscleveland.com/article/20141112/BLOGS03/141119915/parker-hannifin-accepts-the-challenge-of-commercializing-the-indego--41951672-812546791-1442419256=:6352--}\\

\subsubsection{Honda}

\noindent
2 devicesin the walking assist program: 1) Bodyweight Support Assist and
Stride Management Assist (SMA):\\
\url{http://corporate.honda.com/innovation/walk-assist/}\\
\url{http://world.honda.com/Walking-Assist/}\\
\url{http://www.slashgear.com/honda-demo-walking-assist-exoskeleton-for-elderly-disabled-2211295/}\\
\url{http://singularityhub.com/2010/09/13/hondas-exoskeletons-help-you-walk-like-asimo-video/}\\

\subsubsection{Hyundai}

\noindent
Appears to be similar to BLEEX.\\
Article and video: \url{https://forums.xilinx.com/t5/Xcell-Daily-Blog/Hyundai-s-wearable-exoskeleton-restores-personal-mobility-to-the/ba-p/647031}\\
Hyundai's exo project is lead by Dongjin Hyun who did his
PhD with Homayoon Kazerooni at UC Berkeley and a PostDoc with Sangbae
Kim at MIT.

\subsubsection{3D Systems}

\noindent
3D printed exoskeleton. Using Ekso Bionics technology.\\
\url{http://www.dezeen.com/2014/03/05/3d-printed-exoskeleton-helps-paralysed-users-walk/}\\

\subsubsection{ExoAtlet (Russian)}

\noindent
Appears to be similar to BLEEX.\\
\url{http://robohub.org/clinical-trials-begin-for-russias-first-medical-exoskeleton/}\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Academic exoskeletons for assistance and rehabilitation}

\subsubsection{Vanderbilt}

\noindent
Commercialized by Indego/Parker Hannafin/\\
\url{https://en.wikipedia.org/wiki/Vanderbilt_exoskeleton}\\
\url{http://research.vuse.vanderbilt.edu/cim/research_orthosis.html}\\
\url{http://www.dailytech.com/Iron+Man+New+Robotic+Exoskeleton+Helps+Paraplegics+Walk/article29079.htm}\\

\subsubsection{Mindwalker}

\noindent
Key feature is actuation in lateral plane.
Each of the powered joints has a series elastic
actuator, which can deliver 100 Nm torque and 1 kW power.\\
\url{https://mindwalker-project.eu/}\\
\url{http://www.3me.tudelft.nl/en/about-the-faculty/departments/biomechanical-engineering/research/dbl-delft-biorobotics-lab/exoskeleton/}\\
\url{http://www.utwente.nl/ctw/bw/research/projects/MINDWALKER/}\\
\url{http://tnsre.embs.org/2015/03/27/design-control-mindwalker-exoskeleton/}\\

\subsubsection{U. Penn}

\noindent
Student project that got some press.\\
\url{http://www.popsci.com/article/technology/invention-awards-2014-powerful-portable-and-affordable-robotic-exoskeleton}\\

\subsubsection{East China University of Science and Technology}

\noindent
Fully portable hydraulic exoskeleton. Unclear if this is for
assistance or augmentation.\\
Video: \url{http://v.youku.com/v_show/id_XMTg3MTQzNjgw.html}\\
Heng Cao: ``most difficult thing was control bandwidth'', so they
are probably not using high quality valves.
\url{http://mech.ecust.edu.cn/s/131/t/147/6a/07/info27143.htm}

\subsubsection{University of Electronic Science and Technology of China}

\noindent
Prof Hong Cheng did his Posdoc in CMU from 2006-2009 in CS.
He is very famous in China for his exoskeleton work.\\
Chinese Video: \url{http://society.kankanews.com/s/2014-12-17/0016103932.shtml}\\
\url{http://uestcrobot.net/~hcheng/}\\
\url{http://www.uestcrobot.net/?q=node/291}\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Government Programs}

\subsubsection{TALOS}

\noindent
We assume you all are familiar with this program.

\subsubsection{DARPA Warrior Web}

\noindent
We assume you all are familiar with this program.

\subsubsection{Japan: NEDO}

\noindent
There was a Japanese government (NEDO) oriented project, ``Robotic
Devices for Nursing Care Project'', that intended to boost the
practical nursing care robots development, including exoskeleton.\\
\url{http://robotcare.jp/?lang=en}\\
\url{http://robotcare.jp/?page_id=29&lang=en}\\
They made a center to test the safety for robots including exoskeletons.\\
\url{http://www.rtnet-biz.jp/rtsic/}\\

\subsubsection{EU: BALANCE program}

\noindent
\url{http://balance-fp7.eu/}\\
\url{http://cordis.europa.eu/project/rcn/106854_en.html}\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%